{\em pyg4ometry}, a python library for the creation of Monte Carlo particle transport geometries

### 2019-02-18 paper/pyg4ometry.tex
Filling out each section

%% main text
\section{Introduction} \label{sec:introduction}
There are numerous different software codes to simulate the passage of particles through material, so called particle transport (PT) programmes, these 
include MCMP, FLUKA, Geant3 and Geant4. Each code either has a particular specialism or target user community.  PT Monte Carlo simulations have 
diverse uses, shielding calculations for radiological protection, detector performance, medical imaging and therapy, space radiation environment simulation 
are some examples. A fundamental requirement of all of the codes is to supply a computer readable description of the physical geometry that the particles 
are passing through.  The creation of geometry files is typically a very time consuming activity, the simulation validity and performance is directly dependent 
on the quality of the geometry. The requirement of a  geometry system that allows the conversion between files prepared for different codes will enable cross 
checks of the physics processes in particle transport codes. Parallel to the creation of geometry files for PT programs, usually computer assisted design (CAD) 
files exist for systems which need to be simulated in PT programmes. The fundamental geometric representations in CAD files is typically not amenable for PT 
programmes.  For these reasons it is advantageous to create a software tool that allows particle transport code users to rapidly develop geometry files, convert between 
common PT programme geometry formats and finally import CAD models. This paper describes such geometry creation and conversion package 
called {\em pyg4ometry}, implemented in {\em Python} and internally based on the Geant4 geometry description. The main features of  {\em pyg4ometry} 
are a python scripting API to rapidly design parametrised geometry, conversion to and from  FLUKA geometry descriptions, conversion from CAD formats 
(STEP and XXX) based on FreeCAD-OpenCascade and powerful geometry visualisation tools based on VTK.

@@

\subsection{Geant Detector Mark-up Language} 
To exchange geometry descriptions Geant Detector Mark-up Language (GDML) was developed. Geant4 and ROOT can read and write GDML format and it is commonly
used as an exchange format for Geant4 geometries. 
\begin{table}
\centering
\begin{tabular}{|c|c|c|} \hline
GDML tag 		&  Child tags				& pyg4ometry class 			\\ \hline
Defines 			&						&						\\
Solids 			&						&						\\
Materials 			&						&						\\
Structure			&						&						\\ \hline
\end{tabular}
\caption{GDML tags and corresponding {\it pyg4ometry} classes.}
\end{table}

\section{{\em pyg4ometry} design and layout}
The core of {\em pyg4ometry} consists of python classes that mimic Geant4 solids, logical, physical volume objects, GDML parameters material classes.
The constructors of the python classes are kept as close to the original Geant4 C++ implementation so that {\it pyg4ometry} users do not have to learn 
a new API. Geometry construction in python proceeds in a way which is very similar to geometry construction in Geant4. A user relatively familiar with Geant4
should be able to start creating geometry in pyg4ometry immediately.

@@

\subsection{Visualisation}
A  geometry hierarchy can be viewed using the popular Visualisation Toolkit (VTK). No separate scene tree is required as the Geant4 volume hierarchy is sufficient 
to place the meshes associated with each physical volume. The python version of Geant4's physical volume is also used to store visualisation attributes like the solids 
color, surface or wireframe representation and visibility.

@@

\subsection{Input and output} 
For each input format available in {\it pyg4ometry} a dedicated {\it Reader} class is implemented, so \verb|gdml.Reader|, \verb|stl.Reader|, \verb|fluka.Reader| 
and \verb|freecad.Reader|. Each reader constructs the appropriate Geant4 classes and

@@

\section{Conclusions}
The authors believe that tools to quickly create, either ab-initio or by conversion geometry usable in particle transport Monte Carlo programmes will save significant 
amounts of user effort and ultimately yield more accurate simulations. {\it pyg4ometry} is a relatively complete implementation of a geometry creation tool, whilst heavily 
internally based on Geant4 and GDML can have utility for users of all PT codes. {\it pyg4ometry} can clearly be extended to other formats or applications. As for applications,
examples include  the  generation triangular mesh structures for GPU accelerated photon tracking in liquid noble  dark matter detectors and geometries for event visualisation.

>>>>

%% main text
\section{Introduction} \label{sec:introduction}
There are numerous different software codes to simulate the passage of particles through material, so called particle transport (PT) programmes, these 
include MCMP, FLUKA, Geant3 and Geant4. Each code either has a particular specialism or target user community.  PT Monte Carlo simulations have 
diverse uses, shielding calculations for radiological protection, detector performance, medical imaging and therapy, space radiation environment simulation 
are some examples. A fundamental requirement of all of the codes is to supply a computer readable description of the physical geometry that the particles 
are passing through.  The creation of geometry files is typically a very time consuming activity, the simulation validity and performance is directly dependent 
on the quality of the geometry. The requirement of a  geometry system that allows the conversion between files prepared for different codes will enable cross 
checks of the physics processes in particle transport codes. Parallel to the creation of geometry files for PT programs, usually computer assisted design (CAD) 
files exist for systems which need to be simulated in PT programmes. The fundamental geometric representations in CAD files is typically not amenable for PT 
programmes.  For these reasons it is advantageous to create a software tool that allows particle transport code users to rapidly develop geometry files, convert between 
common PT programme geometry formats and finally import CAD models. This paper describes such geometry creation and conversion package 
called {\em pyg4ometry}, implemented in {\em Python} and internally based on the Geant4 geometry description. The main features of  {\em pyg4ometry} 
are a python scripting API to rapidly design parametrised geometry, conversion to and from  FLUKA geometry descriptions, conversion from CAD formats 
(STEP and IGES) based on FreeCAD-OpenCascade and powerful geometry visualisation tools based on VTK.

@@

\subsection{Geant Detector Mark-up Language} 
To exchange geometry descriptions Geant Detector Mark-up Language (GDML) was developed. Geant4 and ROOT can read and write GDML format and it is commonly
used as an exchange format for Geant4 geometries. 
\begin{table}[hbt!]
\centering
\begin{tabular}{|c|c|c|} \hline
GDML tag 		&  Child tags				& pyg4ometry class 			\\ \hline
Defines 			&						&						\\
Solids 			&						&						\\
Materials 			&						&						\\
Structure			&						&						\\ \hline
\end{tabular}
\caption{GDML tags and corresponding {\it pyg4ometry} classes.}
\end{table}

\section{{\em pyg4ometry} design and layout}
The core of {\em pyg4ometry} consists of python classes that mimic Geant4 solids, logical, physical volume objects, GDML parameters and material classes.
The constructors of the python classes are kept as close to the original Geant4 C++ implementation so that {\it pyg4ometry} users do not have to learn 
a new API. Geometry construction in python proceeds in a way which is very similar to geometry construction in Geant4. A user relatively familiar with Geant4
should be able to start creating geometry in pyg4ometry immediately. In the following sections novel or important developments in {\it pyg4ometry} are described.

@@

@@

\subsection{Input and output} 
For each input format available in {\it pyg4ometry} a dedicated {\it Reader} class is implemented, so \verb|gdml.Reader|, \verb|stl.Reader|, \verb|fluka.Reader| 
and \verb|freecad.Reader|. Each reader constructs the appropriate Geant4 classes and provides a registry which can be used or manipulated by the user. 
Output consists of taking the registry and writing the appropriate GDML.

\subsection{Overlap detection}
All PT transport codes can not handle spatial overlap between two geometric objects and likely are to have ill defined behaviour when encountering such a 
situation.  A key feature of {\it pyg4ometry} is the detection of potential overlaps efficiently, it does this by performing a CSG intersection between solid instances 
and determining if the resulting mesh is empty.

\subsection{Visualisation}
A  geometry hierarchy can be viewed using the popular Visualisation Toolkit (VTK). No separate scene graph is required as the Geant4 volume hierarchy is sufficient 
to place the meshes associated with each physical volume. The python version of Geant4's physical volume is also used to store visualisation attributes like the solids 
color, surface or wireframe representation and visibility. Overlaps detected in the discretised mesh geometry can be displayed separately to allow a user to remove
the potential overlaps. This is dependent on the accuracy of the mesh generated and is a particular problem with meshes based on curved surfaces.

@@

\section{Conclusions}
The authors believe that tools to quickly create, either ab-initio or by conversion geometry usable in particle transport Monte Carlo programmes will save significant 
amounts of user effort and ultimately yield more accurate simulations. {\it pyg4ometry} is a relatively complete implementation of a geometry creation tool, whilst heavily 
internally based on Geant4 and GDML can have utility for users of all PT codes. {\it pyg4ometry} can clearly be extended to other formats or applications. As for applications,
examples include  the  generation triangular mesh structures for GPU accelerated photon tracking in liquid noble  dark matter detectors and geometries for event visualisation. 
Although not implemented in the current version, a C++ output writer can be quickly implemented to generate geometry which can be compiled into a Geant4 application.
